---
title       : A Linear Regression Model for Infant Birth Weight 
subtitle    : 3.14.15
author      : Mark K. 
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Introduction

- Low birth weight is associated with mortality, reduced growth and cognitive development, and chronic disease in adult life ([see.](https://en.wikipedia.org/wiki/Low_birth_weight))  

- The CDC [reports](http://www.cdc.gov/nchs/fastats/birthweight.htm) **8.0%** of infants born in **2013** were of low birth weight (weight less than 2.5 kg.)  

- I created a shiny app that uses the [birthwt](https://stat.ethz.ch/R-manual/R-patched/library/MASS/html/birthwt.html) data set in the MASS package to examine linear models of risk factors for low birth weight.

--- .class #id 

## Risk Factors for Low Birth Weight  
  
For this project I decided to focus on five risk factors that might explain low birth weight (bwt: kg.)

- Mother's age (age)
- Mother's weight at last menstrual period (lwt: lbs)
- Mother's race (race: white, black, other) 
- Smoking status during pregnancy (smoke: yes, no)
- History of hypertension (ht: yes, no)  
  
You can create different models with combinations of these risk factors using my [app.](https://mark-ko.shinyapps.io/myApp/
)

---

## Results  
  
The models showed some statistically significant linear regression coefficients, for example: 
```{r models, echo=FALSE}
library(MASS)
fit <- lm(bwt ~ lwt + smoke, data = birthwt)
round(summary(fit)$coef[,c(1,4)], 5)
```
But, the adjusted R-squared values were low for this particular model: 
```{r adjr2, echo=FALSE}
round(summary(fit)$adj, 3)
```
 

---  

## Conclusion  

  
Combinations of the covariates will produce linear regression coefficients with low p-values, but the models have low adjusted R-squared. This is likely due to the variability in the covariates.  
```{r plot, echo=FALSE, fig.height = 4}
plot(fit, which = 1)
```
  
These simple linear models show risk factors, but are <strong>not adequate for prediction.</strong> There could be interactions between risk factors not modeled, or important risk factors may not be included in the data set.

